Minimax lower bounds for source localization

### 2017-01-10 paper.tex
Add first draft of introduction.

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{cite}
\usepackage[pdftex]{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{array}
\usepackage{enumitem}
\usepackage{color}
\usepackage{url}
\usepackage[colorlinks=true,allcolors=black,urlcolor=blue]{hyperref}
%\usepackage{caption}
%\usepackage{subcaption}

@@

\textcolor{red}{Introduction needs to be re-written.}

Electroencephalograhy (EEG) is a system used to record the electrical activity
of the brain. It is a non-invasive system, which may use anywhere from around
10 to 250 electrodes placed on the scalp, to sense electric potentials. These
potentials are generated by neuronal activity within the
brain~\cite{Buzsaki2012Origin}. These sources of neural activity -- neurons or
groups of neurons -- are usually modeled as current
dipoles~\cite{Nunez2006Electric}.

Often, it is of clinical or scientific interest to reconstruct brain activity
patterns from EEG sensor measurements. This process is what is broadly referred
to when we speak of the term ``source localization''. In this paper, I restrict
the discussion to that of localizing a \emph{single} dipole, and refer to this
process as ``dipole source localization''. Several algorithms for source
localization have been proposed over the years
(see~\cite{Baillet2001Electromagnetic} for an excellent review). However, until
recently, there has been only one work that has theoretically examined the
fundamental limits of source localization accuracy.

Mosher et.\ al.~\cite{Mosher1993Error} first propounded Cramer-Rao lower bounds
for unbiased source localization algorithms. Their work also forms the basis
for setting up the source localization problem and deriving more universally
applicable bounds in this paper. The Cramer-Rao lower bound is not in itself a
completely satisfactory answer to the question of fundamental limits for source
localization accuracy. This is because Cramer-Rao bounds only apply to unbiased
estimators (or to estimators of \emph{known} bias). In practice, however, many
popular source localization algorithms have been shown to be biased, e.g.\
the Minimum Norm Estimate (MNE)~\cite{Hamalainen1994Interpreting} is known to
bias the solution towards the surface of the head~\cite{Lin2006Assessing}. It
is also known that unbiased estimators can have worse mean-squared-error
performance than biased estimators. So Cramer-Rao bounds are useful only to a
degree, and there may be benefit in deriving bounds that hold for all
estimators.

To this end, Grover~\cite{Grover2016Fundamental} recently proposed a new lower
bound for the error in localizing a single dipole. This bound used techniques
from information theory, for bounding the source localization error by
expressing it in terms of the mutual information across some channel, and then
bounding the mutual information by the capacity of the channel. In deriving
this bound, however, the scalp potential was never sampled; equivalently, it
was assumed that an \emph{infinite} number of sensors was available to sample
the EEG signal at every point on the scalp. Clearly, this is a poor assumption
and could result in a very loose bound, because it allows for more information
to be available at the receiver than otherwise possible. The lower bound also
relies on certain relaxations that allow the dipole to distribute its power in
non-physical ways in frequency domain. This once again makes the bound loose by
assigning the channel a higher capacity than what it could otherwise have.

To address the shortcomings of the Cramer-Rao and information theoretic lower
bounds, I apply two minimax methods to the problem of computing a lower bound
for localization error in the idealized EEG dipole source localization problem.
I briefly review two minimax methods -- Le Cam's, and Fano's local
method~\cite{Duchi2015Information} -- and describe my approach in deriving the
minimax lower bound using each of these for spherical and one-dimensional head
models.

@@

\section{Minimax lower bounds for the one-dimensional model}

@@

\section{Extensions}

\subsection{Linear domain (in place of a circle)}

@@

Another simple extension takes our problem from a one-dimensional setting to a
$d$-dimensional setting. Here, we would have $\v\theta \in [0, T)^d$, and the
$\ell_2$ loss function, $\Phiorho(\v\theta, \vhat\theta) = \norm{\v\theta -
\vhat\theta}^2$. Then, using Assouad's method~\cite{Duchi2015Information}, we
can split the minimax risk along each of the $d$ dimensions, and bound each
separately using Le Cam's method.

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{cite}
\usepackage[pdftex]{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{array}
\usepackage{enumitem}
\usepackage{color}
\usepackage{url}
\usepackage[colorlinks=true,allcolors=black,urlcolor=blue]{hyperref}

\urlstyle{tt}

%\usepackage{caption}
%\usepackage{subcaption}

@@

%- The source localization problem is one that occurs in several fields
%  - Brain activity measurement, in the form of localizing brain activity
%    - Possible applications are epilepsy focus localization
%    - Neuroscientific experiments - locating the source of activity at certain
%      points in time, for understanding brain function
%  - Need to read Vetterli's papers on acoustic source localization-type things
%  - Need to read up linear inverse problems literature

%- Previous work
%  - In the brain activity localization side of things, lots of algorithms, but
%    not much understanding of bounds (Mosher, Grover).
%    - These have their own issues: CRLB does not apply to all estimators, info
%      theoretic bound does not give scaling with sensors
%  - Other work on minimax bounds and estimators for linear inverse problems
%    concentrate on function estimation (with smoothness constraints on the
%    function being estimated)
%    - e.g. Tsybakov, Efromovich, Ibragimov \& Has'minskii, and references
%      therein
%    - Obviously, bounds for function estimation will look radically different
%      from bounds for a location parameter
%    - Location-parameter based work appears to exist in Ibragimov-Has'minskii,
%      but they treat it as an exercise, without seriously considering
%      extensions, and looking at scaling with number of sensors

%- Cite Fisher's book on circular domain signals somewhere (statistical analysis
%  of circular data)

The source localization problem is one that appears in many fields where it is
of interest to find the position of a source (which emits some kind of signal)
using an array of sensors.

Our principal motivation comes from an application in brain activity
measurement, where one wishes to use modalities such as electroencephalography
(EEG) and magnetoencephalography (MEG) to record the activity inside the
brain~\cite{Baillet2001Electromagnetic}.  For example, current EEG systems use
anywhere between 10 and 250 sensors (electrodes), placed on the scalp, to sense
electric potentials produced by neuronal activity within the
brain~\cite{Nunez2006Electric}.

Recent developments~\cite{Grover2016Information} have suggested that source
localization can benefit greatly from an increase in sensor density. This
conclusion is based primarily on a Nyquist rate analysis, and on
information-theoretic bounds (which are possibly very loose) on the accuracy of
source localization algorithms. These bounds, first derived
in~\cite{Grover2016Fundamental}, do not capture scaling in number of sensors,
however, as the derivation effectively assumes the presence of an infinite
number of sensors to derive the bound.

Earlier work by Mosher et al.~\cite{Mosher1993Error}, which seeks to address
this issue by deriving Cramer-Rao lower bounds, is also not completely
satisfactory because these bounds do not apply to biased estimators. Commonly
used source localization algorithms (such as the Minimum Norm
Estimate~\cite{Hamalainen1994Interpreting}) on the other hand, are known to be
biased~\cite{Lin2006Assessing}.

This paper seeks to address the shortcomings of earlier methods by deriving
minimax lower bounds for the source localization problem, which apply to
\emph{all} estimators (biased \emph{and} unbiased), and which can give a
scaling in terms of the number of sensors used (for a fixed sensor placement
configuration). Computing the lower bound for the ``real brain model'' (where
the complete geometry of the brain surface is accounted for) is rendered very
difficult since analytical solutions for the EEG signal, given the brain
activity, do not exist. ``Spherical head
models''~\cite{Nunez2006Electric,Grover2016Information} \emph{do} have
analytical solutions, but are still not easily amenable to analytical lower
bounds because the spherical surface does not permit uniform
sampling~\cite{ImpossibilityOfSphereSampling}, and because the solution is
expressed in terms of special functions. We therefore restrict our analysis to
a source localization problem on a one-dimensional domain (to be described in
detail in Section~\ref{sec:source-localization}).

The one-dimensional toy problem is an effective tool for understanding bounds
on source localization accuracy in other settings as well, such as
\textcolor{red}{<insert other applications>}. A vast literature on linear
inverse problems (refer~\cite{Bal2012Introduction} for an introduction to this
field) and deconvolution algorithms exists, and has addressed the reduced
one-dimensional problem in broad settings
(see~\cite{Cavalier2002Sharp,Efromovich1997Robust,Ibragimov1981Statistical} and
references therein). However, our setting and interpretation appears to be
unique, since most prior work focuses on recovering a whole function (given
certain smoothness constraints), rather than locating a point source. Work that
does address point sources, to our knowledge, does not address scaling in the
number of sensors, since their problem setting and application is very
different.

Hence, we believe that this paper is the first to give lower bounds on source
localization error (measured in an $ell_2$-distance sense) in estimating the
location of a one-dimensional point source, which is observed by sensors
through a diffusive medium (treated as a low-pass filter), and corrupted by
additive Gaussian noise. We describe the problem setting in detail in
Section~\ref{sec:source-localization}, describe the minimax technique used and
give our main result in Section~\ref{sec:minimax-lower-bounds}, describe a few
straightforward extensions in Section~\ref{sec:extensions}, and conclude with a
discussion on implications for sensor placement and future work in
Section~\ref{sec:discussion}.

%\textcolor{red}{Introduction needs to be re-written.}

%Electroencephalograhy (EEG) is a system used to record the electrical activity
%of the brain. It is a non-invasive system, which may use anywhere from around
%10 to 250 electrodes placed on the scalp, to sense electric potentials. These
%potentials are generated by neuronal activity within the
%brain~\cite{Buzsaki2012Origin}. These sources of neural activity -- neurons or
%groups of neurons -- are usually modeled as current
%dipoles~\cite{Nunez2006Electric}.

%Often, it is of clinical or scientific interest to reconstruct brain activity
%patterns from EEG sensor measurements. This process is what is broadly referred
%to when we speak of the term ``source localization''. In this paper, I restrict
%the discussion to that of localizing a \emph{single} dipole, and refer to this
%process as ``dipole source localization''. Several algorithms for source
%localization have been proposed over the years
%(see~\cite{Baillet2001Electromagnetic} for an excellent review). However, until
%recently, there has been only one work that has theoretically examined the
%fundamental limits of source localization accuracy.

%Mosher et.\ al.~\cite{Mosher1993Error} first propounded Cramer-Rao lower bounds
%for unbiased source localization algorithms. Their work also forms the basis
%for setting up the source localization problem and deriving more universally
%applicable bounds in this paper. The Cramer-Rao lower bound is not in itself a
%completely satisfactory answer to the question of fundamental limits for source
%localization accuracy. This is because Cramer-Rao bounds only apply to unbiased
%estimators (or to estimators of \emph{known} bias). In practice, however, many
%popular source localization algorithms have been shown to be biased, e.g.\
%the Minimum Norm Estimate (MNE)~\cite{Hamalainen1994Interpreting} is known to
%bias the solution towards the surface of the head~\cite{Lin2006Assessing}. It
%is also known that unbiased estimators can have worse mean-squared-error
%performance than biased estimators. So Cramer-Rao bounds are useful only to a
%degree, and there may be benefit in deriving bounds that hold for all
%estimators.

%To this end, Grover~\cite{Grover2016Fundamental} recently proposed a new lower
%bound for the error in localizing a single dipole. This bound used techniques
%from information theory, for bounding the source localization error by
%expressing it in terms of the mutual information across some channel, and then
%bounding the mutual information by the capacity of the channel. In deriving
%this bound, however, the scalp potential was never sampled; equivalently, it
%was assumed that an \emph{infinite} number of sensors was available to sample
%the EEG signal at every point on the scalp. Clearly, this is a poor assumption
%and could result in a very loose bound, because it allows for more information
%to be available at the receiver than otherwise possible. The lower bound also
%relies on certain relaxations that allow the dipole to distribute its power in
%non-physical ways in frequency domain. This once again makes the bound loose by
%assigning the channel a higher capacity than what it could otherwise have.

%To address the shortcomings of the Cramer-Rao and information theoretic lower
%bounds, I apply two minimax methods to the problem of computing a lower bound
%for localization error in the idealized EEG dipole source localization problem.
%I briefly review two minimax methods -- Le Cam's, and Fano's local
%method~\cite{Duchi2015Information} -- and describe my approach in deriving the
%minimax lower bound using each of these for spherical and one-dimensional head
%models.

@@

\section{Minimax lower bounds for the one-dimensional model}
\label{sec:minimax-lower-bounds}

@@

\section{Extensions}
\label{sec:extensions}

\subsection{From a circular (periodic) domain to a linear (aperiodic) domain}

@@

Another simple extension takes our problem from a one-dimensional setting to a
$d$-dimensional setting. Here, we would have $\v\theta \in [0, T)^d$, and the
$\ell_2$ loss function, $\Phiorho(\v\theta, \vhat\theta) = \norm{\v\theta -
\vhat\theta}^2$. Then, using Assouad's lemma~\cite{Tsybakov} in the manner
described in~\cite{Duchi2015Information}, we can split the minimax risk
along each of the $d$ dimensions, and bound each separately using Le Cam's
method.
